---
title: "Research questions"
output: html_document
author: "Tjeu Groothuijsen"
bibliography: references.bib
---


```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled yesterday or something

<br>

----
  
Something that has kept me busy for a few years now are the various specialties and
distinctions in sociology, specifically in the Netherlands. I am hesitant to make claims
about my observations here because I know some of you reading this know far more about it
than I do, but during the third year of my bachelor I got the impression that there is a
clearly distinct group called 'analytical sociology' (or various other names depending on
who you ask) which is quite well represented in the Netherlands. This group their own
professional association and looks to be linked to a graduate school called 'ICS'.

Recently I read an article that described the development of sociology in the Netherlands
which also went into some depth regarding the differences between two prominent graduate
schools, ICS and AISSR [@Haan]. Apparently my view that there are topical and
methodological differences between these schools is to some extent correct, with them
having (or having had) distinct approaches to sociology.

This lead me to question whether the graduates of the two schools formed separate groups,
or if they perhaps integrate and start collaborations regardless of where they studied. 
I think if they stay separated it might lead to clear clustering in the network, so I
found some articles which study the extent of clustering in science collaboration.

Moody studied a social science network and found no clustering [@Moody]. Studies by
Newman and Tomassini and Luthi, however, both found high clustering [@Newman][@Tomas].

So does clustering occur in our data?

And my real question: is the clustering caused by grad school affiliation?

  
